---
title: 'HW #2'
author: "Ashley Kang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
This problem focuses on NYC Transit data on entrances and exits of each subway station in NYC

### Reading and cleaning data

* Read and clean the data
* Retain line, station name, station latitude/longitude, routes served, entry, vending, entrance type, and ADA compliance. 
* Convert the entry variable from character to a logical variable 
```{r read_clean_1}
nyc_transit = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select("line", "station_name", "station_latitude", "station_longitude", "route1":"route11", "entry", "vending", "entrance_type", "ada") %>% 
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
```

The `nyc_transit` dataset contains **`r nrow(nyc_transit)` rows** and **`r ncol(nyc_transit)` columns**. This dataset consists of character variables such as `line` and `station_name` that describe the name of the station and where the line runs and coordinate variables (numeric) such as `station_latitutde` and `station_longitude` that describe the location of the entrance/exit of each station. Variables `route1` - `route11` (some character, some integers) tell us which trains (e.g. A, B, Q, 4, etc.) run through each station. Other variables including `vending` (character), `entrance_type`(character), `ada` (logical) and `entry` - which was converted from a character variable to a logical variable - were also included in the dataset.  

The dataset was read in and `janitor::clean_names()` was utilized to clean the variable names. Next, `select` was implemented to select the above variables. Lastly, the `entry` variable was converted from a character to a logical variable by using `mutate` and `recode`. Yes, the data are tidy.

#### Answer the following questions:
##### How many distinct stations are there?
```{r distinct_stations_1}
dist_num_station = (distinct(nyc_transit, station_name, line))
```
There are **`r nrow(dist_num_station)`** distinct stations.

##### How many stations are ADA compliant?
```{r distinct_ada_1}
ada_station = distinct(nyc_transit, station_name, line, ada) %>%
  filter(ada == TRUE)
```
There are **`r nrow(ada_station)`** distinct ADA compliant stations.

##### What proportion of station entrances/exits without vending allow entrance?
```{r no_vending_allow_entrance_1}
no_vending = filter(nyc_transit, vending == "NO")
allow_entrance = filter(no_vending, entry == TRUE)
```

The proportion of station entrances/exits without vending that allow entrance is **`r nrow(allow_entrance)/nrow(no_vending)`** (= **37.7%**).

### Reformatting data
Reformat so that route number and route name are distinct variables
```{r reformatting_1}
route_reformat = nyc_transit %>%
  gather(key = route_number, value = route_name, route1:route11)
```

##### How many distinct stations serve the A train?
```{r distinct_a_1}
dist_a_station = distinct(route_reformat, line, station_name, .keep_all = TRUE) %>%
  filter(route_name == "A")
```
There are **`r nrow(dist_a_station)`** distinct stations that serve the A train.

##### Of the stations that serve the A train, how many are ADA compliant?
```{r}
dist_ada_a_station = filter(dist_a_station, ada == TRUE)
```

Of the distinct stations that serve the A train, there are **`r  nrow(dist_ada_a_station)`** ADA compliant stations.

## Problem 2
This problem uses the Mr. Trash Wheel dataset

### Reading and cleaning data

* Specify the sheet in the Excel file and to omit columns containing notes 
* Use reasonable variable names
* Omit rows that do not include dumpster-specific data
* Round the number of sports balls to the nearest integer and converts the result to an integer variable

```{r read_clean_2}
trashwheel = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "Mr. Trash Wheel", range = cellranger::cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  mutate(sports_balls = round(sports_balls)) %>%
  mutate(sports_balls = as.integer(sports_balls))
```

#### Read and clean precipitation data for 2016 and 2017

Omit rows without precipitation data and add a variable year

##### Precipitation data (2016)
```{r read_clean_precipit_2016_2}
precipit_16 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation", range = cellranger::cell_cols("A:B")) %>%
  janitor::clean_names() 

names(precipit_16) = precipit_16[1,]  
precipit_16 = precipit_16[-c(1),]

precipit_2016 = precipit_16 %>% 
  filter(!is.na(Total)) %>%
  filter(!is.na(Month)) %>%
  mutate(Year = "2016")
```

##### Precipitation data (2017)
```{r read_clean_precipit_2017_2}
precipit_17 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation", range = cellranger::cell_cols("A:B")) %>%
  janitor::clean_names() 

names(precipit_17) = precipit_17[1,]  
precipit_17 = precipit_17[-c(1),]

precipit_2017 = precipit_17 %>% 
  filter(!is.na(Total)) %>%
  filter(!is.na(Month)) %>%
  mutate(Year = "2017")
```

##### Combine datasets and convert month to a character variable
```{r read_clean_precipit_combined_2}
precipit_combined = left_join(precipit_2016, precipit_2017, by = "Month") %>%
  rename(Total_2016 = Total.x, Total_2017 = Total.y) %>%
  select(-Year.x, -Year.y) %>%
  mutate(Month = month.name)
```

